\documentclass{article}
\include{sem}
\include{sentences-1file}
\begin{document}

\section{Anatomy of a relative reading}

\sentence{John$_F$ climbed the highest mountain. [relative]}

\begin{itemize}
\item Contrast set: John and his focus-alternatives
\item Association relation: $x$ climbed (mountain) $y$
\item Measured entities: mountain 1, mountain 2, etc.
\item Measure: $y$ is $d$-high 
\item Domain of measured entities: mountain
\end{itemize}

\sentence{John climbed the highest mountain on Tuesday$_F$ [relative]}
\begin{itemize}
\item Contrast set: Tuesday and its focus alternatives (days)
\item Association relation: John climbed (mountain) $y$ on $x$
\item Measured entities: mountain 1, mountain 2, etc.
\item Measure: $y$ is $d$-high
\item Domain of measured entities: mountain
\end{itemize}


\sentence{John climbed the highest mountain. [absolute]}

\begin{itemize}
\item Contrast set: mountain 1, mountain 2, etc.
\item Association relation: $x=y$
\item Measured entities: mountain 1, mountain 2, etc.
\item Measure: $y$ is $d$-high
\item Domain of measured entities: mountain
\end{itemize}


In general, elements of the contrast set are related to elements of
the measured entities via the association relation, and the measured
entities are members of the domain.

Suppose you are a computer asked to identify the pieces of a sentence
containing a superlative. Then you can follow the following procedure.
If there is no focus, then you have an absolute reading:
\begin{itemize}
\item The association relation is identity.
\item The contrast set is given by the head noun.
%the set of measured entities.
\end{itemize}
In the presence of focus,  you can have a relative or an absolute
reading. For the relative reading:
\begin{itemize}
\item The association relation can be derived by
abstracting over the focus and the noun phrase (more generally:
smallest maximal projection) containing the superlative
phrase. %cf. Farkas & Kiss argument against contextually given assoc relns
\item The contrast set is (a contextually restricted
  subset of) the focus alternatives.
\item The measured entities are the members of the range of the
  association relation (the set of $y$s such that $xRy$).
\end{itemize}
For the absolute reading: follow the procedure for the case where
there is no focus.


% to interpret a superlative sentence: 

% first figure out ass'n rel'n via focus;
% that gives you c.s. + m.e.; (by abstracting over their positions -
% focus for the contrast set, and superlative phrase for the m.e.) 
%      BUT if there is no focus, then a.r. defaults to =;
%      = can't give you c.s. or m.e.; therefore
%            you default to the NOUN's extension (some subset thereof)
%            to give you both c.s. + m.e. and you then get an absolute
%            reading;
%            BUT if you cannot default to the noun b/c there is no noun
%            (as is the case w adverbials and predicative),
%                 THEN you cannot use = as a.r.; 
%                 AND therefore cannot get an absolute reading
           





\section{Adverbial superlatives}

\paragraph{Question:} Do adverbial superlatives have both absolute and relative readings?

\paragraph{Answer:} They certainly seem have relative readings.

\sentence{John$_F$ drove (the) fastest.\label{drove-the-fastest}}

This looks like a relative reading either like this (\textbf{degree hypothesis}):
\begin{itemize}
\item Contrast set: John and his focus-alternatives.
\item Association relation: $x$ ran $y$-fast
\item Measured entities: running speed 1, running speed 2, etc.
\item Measure: $y$ is $d$(-great)
\item Domain of measured entities: speeds
\end{itemize}
or like this (\textbf{event hypothesis}):
\begin{itemize}
\item Contrast set: John and his focus-alternatives
\item Association relation: $x$ is the agent of $y$
\item Measured entities: running event 1, running event 2, etc.
\item Measure: $y$ is $d$-fast
\item Domain of measured entities: running events
\end{itemize}
What would an absolute reading look like? The association relation
would be identity, so the contrast set is identical to the set of
measured entities. 
\begin{itemize}
\item \textbf{Absolute reading (degree hypothesis)}\\
John drove $y$-fast, and $y$ is greater in magnitude than all other
speeds in the context.
\item \textbf{Absolute reading (event hypothesis)}\\
John was the agent of driving event $y$, and $y$ is faster than all
other driving events in the context.
\end{itemize}
Prediction of degree hypothesis: \ref{tutor-fastest} should be
possible, and mean the same thing as \ref{tutor-the-fastest-speed}.

\sentence{John has a private driving tutor. The tutor says that John
  can choose to drive 40mph, 50mph, or 60mph. John was unsure which
  to choose, but in the end...
\begin{subsentences}
\item{...he drove the fastest speed.\label{tutor-the-fastest-speed}}
\item{*...he drove fastest.\label{tutor-fastest}}
\item{?...he drove the fastest.\label{tutor-the-fastest}}
\item{...of the given speeds, he drove the
    fastest.\label{tutor-the-fastest-of}}
\end{subsentences}}

Another prediction of the degree hypothesis, assuming an absolute
reading cannot independently be ruled out, is that \ref{j+s-fastest}
should have a non-contradictory reading, where John and Susie both
drove the fastest of the contextually available speed.
%
\sentence{
\begin{subsentences}
\item{*John drove fastest. Susie did too.\label{j+s-fastest}}
\item{?John drove the fastest. Susie did too.\label{j+s-the-fastest}}
\item{Of all the speeds, John drove *(the) fastest. Susie did too.}
\end{subsentences}}

The event hypothesis does not predict \ref{tutor-fastest} to be
felicitous on an absolute reading, because there is only one driving
events in the context.

Compositional derivation:

\sentence{\textit{fast} $\leadsto \lambda e \lambda d \ldot \const{fast}(e,d)$}
\sentence{\textit{-est}$_\theta$$\leadsto \lambda C_{\<e,t\>}\lambda
  G_{\<d,\<e,t\>\>}\lambda x_\tau$\\$ \exists d [ G(x,d) \land \forall x'
  [[C(x') \land x'\neq x] \rightarrow \neg G(x',d)] ]$\\
`there is a degree to which $x$ is $G$ and no distinct $x'$ in $C$ is $G$'}

\sentence{\textit{fastest}$_C$ $\leadsto \lambda x \ldot \exists d [ \const{fast}(x,d) \land \forall x'
  [[C(x') \land x'\neq x] \rightarrow \neg \textsc{fast}(x',d)] ] $}

\sentence{\textit{drove} $\leadsto \lambda e \ldot \const{driving}(e)$}

\sentence{\textit{drove fastest} $\leadsto \lambda x \ldot
  \const{driving}(x) \land  \exists d [ \const{fast}(x,d) \land \forall x'
  [[C(x') \land x'\neq x] \rightarrow \neg \textsc{fast}(x',d)] ] $}

\sentence{\textit{$\exists$ [\textsc{ag} John] drove fastest} \\
 $\leadsto\exists x \ldot \const{ag}(x,\const{j}) \land
  \const{driving}(x) \land  \exists d [ \const{fast}(x,d) \land \forall x'
  [[C(x') \land x'\neq x] \rightarrow \neg \textsc{fast}(x',d)] ] $
}

%%%%



% to interpret a superlative sentence: 

% first figure out ass'n rel'n via focus;
% that gives you c.s. + m.e.; (by abstracting over their positions -
% focus for the contrast set, and superlative phrase for the m.e.) 
%      BUT if there is no focus, then a.r. defaults to =;
%      = can't give you c.s. or m.e.; therefore
%            you default to the NOUN's extension (some subset thereof)
%            to give you both c.s. + m.e. and you then get an absolute
%            reading;
%            BUT if you cannot default to the noun b/c there is no noun
%            (as is the case w adverbials and predicative),
%                 THEN you cannot use = as a.r.; 
%                 AND therefore cannot get an absolute reading
  
\begin{verbatim}
Getting a relative reading for John_F drove (the) fastest with the recipe:

John is focused. Abstract over ``drive fastest'' to get other slot.
Association relation = x is agent of (driving event) e.
Therefore c.s. = [[John]]^F
m.e.s = driving event 1, 2, 3... (where agent is among c.s.)
measure (type <d,vt>) = e is d-fast

==...==> There is a driving event e that John is the agent of
and for all distinct elements of the contrast set, 
the driving events that they are the agent of are not as fast as e.

Getting a relative reading for John_F climbed (the) highest mountain with the recipe:

John is focused. Abstract over ``the highest mountain'' to get other slot.
Association relation = x climbed y.
Therefore c.s. = [[John]]^F
m.e.s = climbed mountains 1, 2, 3... (where climber is among c.s.)
measure (type <d,et>) = x is d-tall

==...==> There is a mountain x that John climbed
and for all distinct elements of the contrast set, 
the mountains that they climbed are not as tall as x.

TODO: Show that the recipe doesn't work for absolute readings of adverbials.
\end{verbatim}

% A proportional reading of an amount superlative would be:

% \sentence{John mostly laughed.}

% \sentence{*John leastly laughed.}

% Consonant with general observation that negative amount superlatives ({\em least}, {\em fewest}) do not have proportional (smallest part of partition) readings.

% \paragraph{Syntactic reststriction}

% \sentence{*John (the) fastest drove.}

% \paragraph{Morphological restrictions}

% {\em lastly}, {\em *prettiestly}, {\em *quickestly}...

% Periphrastic: {\em most elegantly}

\end{document}